---
output:
  pdf_document: default
  html_document: default
---
1.
(a) The portfolios are constructed at the end of each June, are the intersections of 10 portfolios formed on size (market equity, ME) and 10 portfolios formed on investment (Inv). The size breakpoints for year t are the NYSE market equity deciles at the end of June of year t. Investment is the change in total assets from the fiscal year ending in year t-2 to the fiscal year ending in t-1, divided by t-2 total assets. The Inv breakpoints are NYSE deciles. The portfolios for July of year t to June of t+1 include all NYSE, AMEX, and NASDAQ stocks for which we have market equity data for June of t and total assets data for t-2 and t-1.

(b) We use portfolios of assets because it reduces the errors-in-variables problem. Cross-sectional regressions specify estimated betas as regressors. If the errors in the estimated betas are somehow correlated across assets then the estimation errors would tend to offset each other when the assets are put together into test portfolios. Thus, using portfolios as test assets allows for more efficient estimates of factor loadings which will make factor risk premia to be estimated more precisely.

(c) The value weighted portfolio refers to the composition based on weights of individual stocks proportional to their market capitalization. In equal weighted portfolio all the assets (or stocks) have the same weight.


2. 
(a)

```{r}
library(data.table)
library(readxl)

AMZN <- fread("AMZN.csv", select=c(1, 5))
prices = AMZN$Close

AMZN = AMZN[-1,]
n <- length(prices);
AMZN$log_returns <- log(prices[-1]/prices[-n])*100

AMZN = AMZN[1:251]
data_capm <- fread("FF.csv", select=c(1, 2, 5))
data_capm <- data_capm[892:1142]

Y1 = AMZN$log_returns

Mkt <- data_capm$'Mkt-RF'+ data_capm$RF
model1 = lm(Y1 ~ Mkt)
summary(model1)
confint(model1, level=0.95)
```

i. 
R = 0.6651 + 1.5002(Mkt)

This model doesn't substract risk free rate from AMZN log returns and market excess returns which whould result in R = 0.7206 + 1.5025(Mkt). However, it doesn't change the conclusion.

ii. 
                 2.5 %   97.5 %
(Intercept) -0.7147282 2.044852
Mkt          1.1955436 1.804931

iii. 
Alpha = 0.6651 and is statistically signifcantly different from 0. Its p-value is 0.343, it is larger than 0.05 or 0.1, therefore we do not reject the null hypothese that the alpha is statistically signifcantly different from 0. This indicates that Amazon stock does not seem to signifcantly outperform or underperform the overall market.


iv.
B = 1.5002

H0: B = 1
H1: B > 1

t = (1.5002 - 1) / 0.1547 = 3.23

t > 1.645 (crtitical t)

.: We reject the null. It is more risky than the market and it will have higher return.


b)
i. 
```{r}
P100 <- fread("100Portf.csv")
P100 = P100[219:698]
P100 = as.matrix(P100)
P100 <- P100[,-1]

data_rf <- fread("FF.csv", select=c(1, 2, 5))
data_rf <- data_rf[663:1142]

y = as.matrix(data_rf$RF)
mkt = as.matrix(data_rf$'Mkt-RF')

Alphas = c()
Betas = c()

for(i in 1:ncol(P100)) {   
    Y2 = P100[, i] - y
    model <- lm(Y2 ~ mkt)
    Alphas[i] <- model$coefficients[1]
    Betas[i] <- model$coefficients[2]
}

AVGs = c()
for(i in 1:ncol(P100)) {   
    avg = sum(P100[, i] - y)/480
    AVGs[i] <- avg
}

model_final <- lm(AVGs ~ Betas)

rf = sum(y)/480
rf
avg_mkt = sum(mkt) /480
avg_mkt

summary(model_final)
plot(model_final$residuals, type="l")

Box.test(model_final$residuals, lag=4, type=c("Ljung-Box"))


```
ii. p-value for intercept is smaller than 0.05, therefore it is statistically significant to 0. The intercept/alpha is 2.3931 with a p-value of 0.00, which is lower than 0.05. This
indicates that we will reject the null hypothesis and conclude it is statistically signifcantly to 0. Thus,the CAPM doesnt not hold. The beta is -1.4733, which is lower than 1 This indicates that it is more risky than the market.

iii. The mean of the residual looks steady at the first part, however there is a # sharp drop towards the end. Variance looks a bit unstable as well.

iv. In step 2, estimated intercept  [2.3931]  doesn't equal to average risk free rate [0.3048333] and estimated coefficient [-1.4733] doesn't equal to (average market return [0.754916] - risk free rate [0.3048333]). Therefore, CAPM model is not suitable. 

3.
(a)
```{r}
P100 <- fread("100Portf.csv")
P100 = P100[219:698]
P100 = as.matrix(P100)
P100 <- P100[,-1]

data_rf2 <- fread("FF.csv")
data_rf2 <- data_rf2[663:1142]

y = as.matrix(data_rf2$RF)
mkt = as.matrix(data_rf2$'Mkt-RF')
SMB = as.matrix(data_rf2$SMB)
HML = as.matrix(data_rf2$HML)

Betas_mkt = c()
Betas_SMB = c()
Betas_HML = c()

for(i in 1:ncol(P100)) {   
    Y2 = P100[, i] - y
    model <- lm(Y2 ~ mkt + SMB + HML)
    Betas_mkt[i] <- model$coefficients[2]
    Betas_SMB[i] <- model$coefficients[3]
    Betas_HML[i] <- model$coefficients[4]
}

AVGs2 = c()
for(i in 1:ncol(P100)) {   
    avg = sum(P100[, i] - y)/480
    AVGs2[i] <- avg
}

model_final2 <- lm(AVGs2 ~ Betas_mkt + Betas_SMB + Betas_HML)



```
```{r}
summary(model_final2)


plot(model_final2$residuals, type="l")
```
```{r}
anova(model_final2)
```


The mean and the variances of the residual look very volatile, especially there is a sharp drop on the index 90.

The intercept/alpha is 1.1142 with a p-value of 0.1515, which is higher than 0.05 or 0.1. This indicates that we will not reject the null hypothesis which is statistically signifanct to 0.  CAPM doesnt not hold.  The beta for makert excess is -0.6712, which is smaller than 1. This indicates that it is less risky than the market.

The intercept/alpha is 1.1142 with a p-value of 0.1515, which is higher than 0.05 or 0.1. This indicates that we will not reject the null hypothesis which is statistically significant to 0. CAPM doesn’t not hold. The beta for market excess is -0.6712, which is smaller than 1. This indicates that it is less risky than the market.

The beta for SMB is 0.1949 with a p-value of 0.182080. Since the p-value is greater than 0.05 or 0.1, it is statistically signifcantly different from  0. Also, an increase by 1 of SMB will result an increase in the excess return by 0.3104. 

The beta for HML is  1.2100 with a p-value of 0.000176. Since the p-value is lower than 0.05 or 0.1, it is statistically signifcantly to 0. Also, an increase by 1 of HML will result in increase in the excess return by 1.2100.



(b)
i. From the graphs below we can conclude that betas are time varying and change during the particular interval.
```{r}
P100 <- fread("100Portf.csv")
P100 = P100[219:698]
P100 = as.matrix(P100)
P100 <- P100[,-1]

data_rf2 <- fread("FF.csv")
data_rf2 <- data_rf2[663:1142]

y = as.matrix(data_rf2$RF)
mkt = as.matrix(data_rf2$'Mkt-RF')
SMB = as.matrix(data_rf2$SMB)
HML = as.matrix(data_rf2$HML)

myfun <- function(data, i_begin, i_end) {
  Res = matrix(0, nrow=0, ncol=4)
  for(i in 1:ncol(data)) {
    Temp = c()
    Y2 = data[i_begin:i_end, i] - y[i_begin:i_end]
    model <- lm(Y2 ~ mkt[i_begin:i_end] + SMB[i_begin:i_end] + HML[i_begin:i_end])
    Temp <- model$coefficients
    Res <- rbind(Res, Temp)
  }
  colnames(Res) <- c("Incpt","Mkt_B","SMB_B","HML_B")
  return(Res)
}

Period1 <- myfun(P100, 1, 96)
Period2 <- myfun(P100, 97, 192)
Period3 <- myfun(P100, 193, 288)
Period4 <- myfun(P100, 289, 384)
Period5 <- myfun(P100, 385, 480)


P100_small <- matrix(0, nrow=0, ncol=4)
P100_small <- rbind(P100_small, Period1[1,], Period2[1,], Period3[1,], Period4[1,], Period5[1,])

df <- as.data.frame(P100_small)

  
```
```{r}
library(ggplot2)
ggplot(data = df, aes(x = as.numeric(row.names(df)))) + 
  geom_line(aes(y = Mkt_B, color='Mkt_B')) + 
  geom_line(aes(y = SMB_B, color='SMB_B')) +
  geom_line(aes(y = HML_B, color='HML_B')) + ggtitle("SMALL LoINV") + labs(y="Betas", x = "Period")

```

```{r}

P100_ME5 <- matrix(0, nrow=0, ncol=4)
P100_ME5 <- rbind(P100_ME5, Period1[50,], Period2[50,], Period3[50,], Period4[50,], Period5[50,])

df1 <- as.data.frame(P100_ME5)

ggplot(data = df1, aes(x = as.numeric(row.names(df)))) + 
  geom_line(aes(y = Mkt_B, color='Mkt_B')) + 
  geom_line(aes(y = SMB_B, color='SMB_B')) +
  geom_line(aes(y = HML_B, color='HML_B')) + ggtitle("ME5 INV10") + labs(y="Betas", x = "Period")
```

```{r}

P100_BIG <- matrix(0, nrow=0, ncol=4)
P100_BIG <- rbind(P100_BIG, Period1[100,], Period2[100,], Period3[100,], Period4[100,], Period5[100,])

df2 <- as.data.frame(P100_BIG)
ggplot(data = df2, aes(x = as.numeric(row.names(df)))) + 
  geom_line(aes(y = Mkt_B, color='Mkt_B')) + 
  geom_line(aes(y = SMB_B, color='SMB_B')) +
  geom_line(aes(y = HML_B, color='HML_B')) + ggtitle("BIG HiINV") + labs(y="Betas", x = "Period")
```

ii. From the graphs below we can conlude that the risk premias are time varying and volatile as well. Some of them seem to show a trend.
```{r}

avgsfun <- function(data, beg_i, end_i) {
  res = c()
  for(i in 1:ncol(data)) {   
      avg = sum(data[beg_i:end_i, i] - y[beg_i:end_i])/96
      res[i] <- avg
  }
  return (res)
}

Period1_avgs <- avgsfun(P100, 1, 96)
Period2_avgs <- avgsfun(P100, 97, 192)
Period3_avgs <- avgsfun(P100, 193, 288)
Period4_avgs <- avgsfun(P100, 289, 384)
Period5_avgs <- avgsfun(P100, 385, 480)


myfun_step2 <- function(avgs, data) {
  res = matrix(0, nrow=0, ncol=4)
  temp = c()
  mod <- lm(avgs ~ data[,2] + data[,3] + data[,4])
  temp <- mod$coefficients
  res <- rbind(res, temp)
  colnames(res) <- c("a","lambda_M","lambda_S","lambda_V")
  return(res)
}

Period1_Lambdas <- myfun_step2(Period1_avgs, Period1)
Period2_Lambdas <- myfun_step2(Period2_avgs, Period2)
Period3_Lambdas <- myfun_step2(Period3_avgs, Period3)
Period4_Lambdas <- myfun_step2(Period4_avgs, Period4)
Period5_Lambdas <- myfun_step2(Period5_avgs, Period5)

step2_matrix = matrix(0, nrow=0, ncol=4)
step2_matrix <- rbind(step2_matrix, Period1_Lambdas, Period2_Lambdas, Period3_Lambdas, Period4_Lambdas, Period5_Lambdas)
rownames(step2_matrix) <- c("1","2","3","4","5")
df_final <- as.data.frame(step2_matrix)
df_final
```
```{r}

ggplot(data = df_final, aes(x = as.numeric(row.names(df_final)))) + 
  geom_line(aes(y = lambda_M, color='lambda_M')) + ggtitle("lambda_M") + labs(y="Betas", x = "Period")
```

```{r}
ggplot(data = df_final, aes(x = as.numeric(row.names(df_final)))) + 
  geom_line(aes(y = lambda_S, color='lambda_S')) + ggtitle("lambda_S") + labs(y="Betas", x = "Period")
```
```{r}
ggplot(data = df_final, aes(x = as.numeric(row.names(df_final)))) + 
  geom_line(aes(y = lambda_V, color='lambda_V')) + ggtitle("lambda_V") + labs(y="Betas", x = "Period")
```

